Supplementary Materials And Methods

### 2021-01-27 Supplementary_fig_tab.Rmd
supplementary files again

## 2.1 Sample dataset analysis in absence of interaction factors.

```{r, echo=FALSE}
# Sample dataset refers to minsamp in ceRNAnetsim package.
data("minsamp")
```

```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Dataset in Steady-state. Inıtial expression levels in minsamp network (*Sample* network in manuscript). The network contains two miRNAs and 6 genes with arbitrarily chosen expression values. Refer to Table S1 for exact expression values", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4",
            title = " ") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)
```

```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Dataset. Expression level of Gene2 is increased from 10,000 to 20,000 in order to demonstrate the calculation steps.", fig.width=5, fig.height=4}
priming_graph(minsamp,
              competing_count = Competing_expression,
              miRNA_count = miRNA_expression) %>%
  update_how("Gene2", 2) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red",
            title = " ") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data after up-regulation of Gene2. A) First response of system to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}

@@

```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression level using interaction parameters in Sample+ network. A) Interaction paramaters between genes and miRNAs in Sample+ are shown on network while expression levels can be found in Table S1. B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters are shown on edges. D) Values on edges indicate repression level. Red values indicate repression level affected by region effect (RE) parameter. E) Total repression on G4 from two miRNAs is calculated by adding repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
knitr::include_graphics("fig_s5.png")

@@

$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient. Numbering on edges match the pair order in Table S1.

## 2.3 Sample+ dataset analysis with interaction factors.

@@

```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A) First response of system to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3rd iteration). Although visualisation looks similar to Figure S4B, current counts of genes are drastically different."}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region) %>%
  update_how("Gene2", 2) %>%
  simulate(cycle = 1) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)

@@

## 2.4 Common target perturbation in Sample+ dataset.

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Log2 transformed expression levels of tumor-specific perturbing nodes in tumor and normal tissue samples of 87 patients."}

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4, fig.height=4, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients."}

@@

## 3.1 *Sample+* dataset

@@

Table: The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in Sample+ dataset, while these factors are not utilized in simulation of Sample dataset.

@@

Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other hand, we were examined other interacion factors based on different studies.

+ [@helwak_mapping_2013; @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
+ Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive]. 
+ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]

@@

Table: Context of miRNA:target pairs supported by High-throughput Experiments. CLEAR-CLiP and CLASH datasets were integrated as described in Section 2 of Supplementary Material and Method.

@@

As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at Table S5.

@@

## 3.5 Comparison of perturbation efficiency between Sample and Sample+ dataset

@@

# 4. Access to code

All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).

>>>>

## 2.1 Sample data set analysis in absence of interaction factors.

```{r, echo=FALSE}
# Sample data set refers to minsamp in ceRNAnetsim package.
data("minsamp")
```

```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Data set in Steady-state. Inıtial expression levels in minsamp network (*Sample* network in manuscript). The network contains two miRNAs and 6 genes with arbitrarily chosen expression values. Refer to Table S1 for exact expression values", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4",
            title = " ") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)
```

```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Data set. Expression level of Gene2 is increased from 10,000 to 20,000 in order to demonstrate the calculation steps.", fig.width=5, fig.height=4}
priming_graph(minsamp,
              competing_count = Competing_expression,
              miRNA_count = miRNA_expression) %>%
  update_how("Gene2", 2) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red",
            title = " ") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data after up-regulation of Gene2. A) First response of network to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}

@@

```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression level (counts) using interaction parameters in Sample+ network. A) Interaction paramaters between genes and miRNAs in Sample+ are shown on network while expression levels can be found in Table S1. B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters (Energy and Seed Type Effect) are shown on edges. D) Values on edges indicate degredation level (couns). Red values indicate degredation level affected by region effect (RE) parameter. E) Total repression on G4 from two miRNAs is calculated by summing repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
knitr::include_graphics("fig_s5.png")

@@

$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type effect; $RE^\prime$, normalized values of region effect parameter; $E^\prime$, normalized values of Energy parameter. Numbering on edges match the pair order in Table S1.

## 2.3 Sample+ data set analysis with interaction factors.

@@

```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A) First response of network to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3rd iteration). Although visualisation looks similar to Figure S4B, current counts of genes are drastically different."}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region) %>%
  update_how("Gene2", 2) %>%
  simulate(cycle = 1) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)

@@

## 2.4 Common target perturbation in Sample+ data set.

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Log2 transformed expression levels of tumor-specific perturbing nodes in tumor and normal tissue samples of 87 patients for Real network."}

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4, fig.height=4, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients for Real data set."}

@@

## 3.1 *Sample+* data set

@@

Table: The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in Sample+ data set, while these factors are not utilized in simulation of Sample data set.

@@

Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other hand, we were examined other interaction parameters based on different studies.

+ [@helwak_mapping_2013; @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
+ Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive]. 
+ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].

@@

Table: Context of miRNA:target pairs supported by High-throughput Experiments. CLEAR-CLiP and CLASH data sets were integrated as described in Section 2 of Supplementary Material and Method.

@@

As a result of simulation a data set, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at Table S5.

@@

## 3.5 Comparison of perturbation efficiency between Sample and Sample+ data set

@@

# 4. Notes and Access to code

+ *Table S7*: refers functional annotations of all highly perturbing genes from simulations of network retrieved from miRTarBase.

+ *Table S8*: refers functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRTarBase.

+ *Table S9*: refers functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRNA:target pairs that are validated by high-throughput experimental studies.

+ All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).

### 2021-01-27 Supplementary_mat_met.Rmd
supplementary files again

Energy values in miRNA:target pairs were represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which were utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we have handled this integrated data set in context of competitor behaviors and functionality of interactions.

@@

Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP data set to show as found in CLASH data set. Same type formatted values added data set as numeric values (shown in Table S2).

@@

The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics. The data sets contain all the chimeric miRNA:target structures found in the medium during the experiment. On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair. We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions. The step was performed as:

>>>>

Energy values in miRNA:target pairs were represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which were utilized in this study. On the other hand, we specified the other interaction factors (seed type and location of binding region on the target) as numeric values based on the previous studies.[@grimson_microrna_2007] had compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we compiled seed types of miRNA:target interactions as numeric values. We also have re-defined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we handled this integrated data set in context of competitor behaviors and functionality of interactions.

@@

Secondly, we arranged seed type parameters, as in *Seed type manipulation* section, of CLEAR-CLiP data set to show as found in CLASH data set. The values added data set as numeric values (shown in Table S2).

@@

The approaches to detect miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics. The data sets contain all readings from the chimeric miRNA:target pairs, isolated at a particular time. On the other hand, it could be said that the reading is performed in a snapshot fashion. Because of that, the methods can provide different chimeric interactions the same miRNA:target pair. We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions. The step was performed as:
